Nail: A practical interface generator for binary formats

### 2014-04-01 paper/future.tex
consistently use a non-breaking space (~) before \cite

A major problem with the current design of Nail is that the design of the grammar dictates the
interna structure of the software. This makes changing grammars or adding Nail to existing software
awkward. One possible solution to this problem would be to implement a concept similar to Relational
Lenses \cite{bohannon2006relational}, which would allow  the data model to be 'seen' by
the rest of the program through an isomorphism. Such an isomorphism would still be much more concise
than two sets of semantic actions, while allowing changes in syntax, alternative representations
and adaption to legacy systems.

>>>>

A major problem with the current design of Nail is that the design of the grammar dictates the
interna structure of the software. This makes changing grammars or adding Nail to existing software
awkward. One possible solution to this problem would be to implement a concept similar to Relational
Lenses~\cite{bohannon2006relational}, which would allow  the data model to be 'seen' by
the rest of the program through an isomorphism. Such an isomorphism would still be much more concise
than two sets of semantic actions, while allowing changes in syntax, alternative representations
and adaption to legacy systems.

### 2014-04-01 paper/intro.tex
consistently use a non-breaking space (~) before \cite

Common wisdom thus says to take special care when preparing code that deals with untrusted input.
However, with currently prevalent software design methods, it is hard to isolate input-handling code
from the rest of the program, as user input is typically passed through the program and processed in
bits and pieces as a 'shotgun parser'\cite{shotgun-parser}.

@@

Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS
\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
  differently, so the code signature sees different bytes at a virtual address than the executable
  that runs.}\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
  files. Signatures are verified with a Java program, but the program is extracted with a C program.
The java program interprets all fields as signed, whereas the C program treats them as unsigned,
allowing one to replace files in a signed archive.}
\cite{saurik-masterkey} code signing and even the
X.509 protocol underlying SSL \cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
runtime environment or static analysis guarantee memory safety.

Finally, many protocol implementations are so complicated that even without bugs, they inadvertently
provide a full Turing-complete execution environment. Examples include X86 page tables
\cite{bangert2013page}, ELF symbols and relocations\cite{shapiro2013weird}. Even though these
protocols were never intended to be that expressive - and consequentially hard to parse, implementation details of their
recognisers - which de facto became a part of the specification and were relied upon by legitimate
inputs - expose unwanted computational power and have unintended side-effects.
In the offensive research community, this has
been generalised into treating a program as a \textit{weird machine}\cite{bratus2011exploit} that
operates on an input, analogous to a virtual machine operating on bytecode.

Proper input recognition has been shown to be an excellent way of eliminating malicious inputs. In
one case, a  PDF parser implemented in Coq could eliminate over $95\%$ of known malicious PDFs\cite{Bogk-PDF}.
However, manually writing parser code does not scale to the number of file formats  and protocols in
existence and might result in parser code tied to one specific application.

\subsection{Parser generators}
%%TODO
Luckily, we can avoid all of these problems at once if we have a single, executable specification of
what exactly a programs input is supposed to look like. In fact, definitive grammars have long
been written for all sorts of protocols and are the standard way of describing text-based protocols.
Parser generators can automatically create parsers for such languages, and with common classes of
grammars such as regular expressions or deterministic context free languages, that parser can be
expressed in a computationally less powerful form, e.g. a finite automaton or a push down automaton,
limiting the weird machine problem and enabling very efficient implementation. \cite{Knuth1965607}

@@

Unfortunately, grammars are seldom used directly to recognise machine-created input. For example,
the security-critical and very well-engineered MIT Kerberos distribution uses parser
generators, but only for handling configuration files. A notable exception is the Mongrel
web server \footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
Ragel \cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
times to achieve better scaleability and design, yet the grammar could be  re-used across all iterations\cite{patterson-citation}.

 Parser generators for binary protocols were first introduced by the Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
 While previous parser generators could also be used to write grammars for binary
 protocols\footnote{Theoretically speaking, the alphabet over which a grammar is an  abstract set,
   so most algorithms work just as well on an alphabet of $\{0,1\}$}, doing so is impractically
 inconvenient. Hammer allows the programmer to specify a grammar in terms of bits and bytes instead
 of characters. Common concerns, such as endianness and bit-packing are transparently hidden by the
 library. Hammer implements grammars as language-integrated parser combinators, an approach
 popularised by Parsec for Haskell\cite{LeijenMeijer:parsec}. The parser combinator style (to our
 knowledge, first  described in \cite{burge1975recursive}) is a natural way of concisely expressing
 top-down grammars\cite{Danielsson:2010:TPC:1863543.1863585} \footnote{For more background on the history of expressing grammars, see Bryan
   Ford's masters thesis\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
  Hammer then constructs a tree of function pointers which
 can be invoked to parse a given input into an abstract syntax tree (AST).

>>>>

Common wisdom thus says to take special care when preparing code that deals with untrusted input.
However, with currently prevalent software design methods, it is hard to isolate input-handling code
from the rest of the program, as user input is typically passed through the program and processed in
bits and pieces as a 'shotgun parser'~\cite{shotgun-parser}.

@@

Such parser differential vulnerabilities were found in various cryptographic systems, including  iOS\footnote{The XNU kernel and the user code-signing verifier interpret executable metadata
  differently, so the code signature sees different bytes at a virtual address than the executable
  that runs.}~\cite{evaders6} and Android\footnote{Android applications are distributed as .zip
  files. Signatures are verified with a Java program, but the program is extracted with a C program.
The java program interprets all fields as signed, whereas the C program treats them as unsigned,
allowing one to replace files in a signed archive.}~\cite{saurik-masterkey} code signing and even the
X.509 protocol underlying SSL~\cite{DBLP:conf/fc/KaminskyPS10}. Such vulnerabilities occur even if the language,
runtime environment or static analysis guarantee memory safety.

Finally, many protocol implementations are so complicated that even without bugs, they inadvertently
provide a full Turing-complete execution environment. Examples include X86 page tables~\cite{bangert2013page}, ELF symbols and relocations~\cite{shapiro2013weird}. Even though these
protocols were never intended to be that expressive - and consequentially hard to parse, implementation details of their
recognisers - which de facto became a part of the specification and were relied upon by legitimate
inputs - expose unwanted computational power and have unintended side-effects.
In the offensive research community, this has
been generalised into treating a program as a \textit{weird machine}~\cite{bratus2011exploit} that
operates on an input, analogous to a virtual machine operating on bytecode.

Proper input recognition has been shown to be an excellent way of eliminating malicious inputs. In
one case, a  PDF parser implemented in Coq could eliminate over $95\%$ of known malicious PDFs~\cite{Bogk-PDF}.
However, manually writing parser code does not scale to the number of file formats  and protocols in
existence and might result in parser code tied to one specific application.

\subsection{Parser generators}
%%TODO
Luckily, we can avoid all of these problems at once if we have a single, executable specification of
what exactly a programs input is supposed to look like. In fact, definitive grammars have long
been written for all sorts of protocols and are the standard way of describing text-based protocols.
Parser generators can automatically create parsers for such languages, and with common classes of
grammars such as regular expressions or deterministic context free languages, that parser can be
expressed in a computationally less powerful form, e.g. a finite automaton or a push down automaton,
limiting the weird machine problem and enabling very efficient implementation~\cite{Knuth1965607}.

@@

Unfortunately, grammars are seldom used directly to recognise machine-created input. For example,
the security-critical and very well-engineered MIT Kerberos distribution uses parser
generators, but only for handling configuration files. A notable exception is the Mongrel
web server \footnote{\url{http://mongrel2.org/}} which uses a grammar for HTTP written in the
Ragel~\cite{ragel-paper} regular expression language. Mongrel was re-written from scratch multiple
times to achieve better scaleability and design, yet the grammar could be  re-used across all iterations~\cite{patterson-citation}.

 Parser generators for binary protocols were first introduced by the Hammer\footnote{\url{http://github.com/upstandinghackers/hammer/}} parser.
 While previous parser generators could also be used to write grammars for binary
 protocols\footnote{Theoretically speaking, the alphabet over which a grammar is an  abstract set,
   so most algorithms work just as well on an alphabet of $\{0,1\}$}, doing so is impractically
 inconvenient. Hammer allows the programmer to specify a grammar in terms of bits and bytes instead
 of characters. Common concerns, such as endianness and bit-packing are transparently hidden by the
 library. Hammer implements grammars as language-integrated parser combinators, an approach
 popularised by Parsec for Haskell~\cite{LeijenMeijer:parsec}. The parser combinator style (to our
 knowledge, first  described in~\cite{burge1975recursive}) is a natural way of concisely expressing
 top-down grammars~\cite{Danielsson:2010:TPC:1863543.1863585}\footnote{For more background on the history of expressing grammars, see Bryan
   Ford's masters thesis~\cite{ford2002packrat}, which also describes the default parsing algorithm used by Hammer} by composing them from one or multiple sub-parsers. 
  Hammer then constructs a tree of function pointers which
 can be invoked to parse a given input into an abstract syntax tree (AST).

### 2014-04-01 paper/relwk.tex
consistently use a non-breaking space (~) before \cite

Generating parsers and generators from an executable specification is the core concept of Interface
Generators, e.g. in CORBA or more recently\cite{varda2008}. However, while interface
generators work very well for existing grammars, they do not allow full control over the format of
the output, so cannot be used to implement legacy protocols.
Very related work has been done at Bell Labs with the PacketTypes system\cite{mccann2000packet},
however PacketTypes works only as a parser and does not support the expressive power of PEGs, but rather
implements a C like structure model enhanced with data-dependent length fields and constraints.

@@

Memory corruption exploits (frequently) rely on placing controlled values at known (or controlled)
memory locations. Exploit countermeasures  like Address-Space Layout Randomisation\cite{pax-aslr},
 heap cookies \cite{heapcookies} or data execution prevention try to make this harder for
 an attacker. Attackers have developed techniques such as heap spraying \cite{heapspray} and heap
 feng shui\cite{fengshui} to help bypass those mitigations.

Nail parsers use two arena allocators\cite{arena}, one for the temporary syntax tree and one for the
result. If Nail were to use the normal system allocator, the attacker could cause two packets to be
parsed at the same time (or, given a sufficiently complicated input, two structure within the same
grammar to be parsed sequentially) such that the (invalid) intermediate results from one parse,  
 If there is a buffer overflow past the
end of one input, it is much harder for the attacker craft a sequence of packets to trick the allocator into
placing a known value next to that input.

>>>>

Generating parsers and generators from an executable specification is the core concept of Interface
Generators, e.g. in CORBA or more recently~\cite{varda2008}. However, while interface
generators work very well for existing grammars, they do not allow full control over the format of
the output, so cannot be used to implement legacy protocols.
Very related work has been done at Bell Labs with the PacketTypes system~\cite{mccann2000packet},
however PacketTypes works only as a parser, not as an output generator and does not support the expressive power of PEGs, but rather
implements a C like structure model enhanced with data-dependent length fields and constraints.

@@

Memory corruption exploits (frequently) rely on placing controlled values at known (or controlled)
memory locations. Exploit countermeasures  like Address-Space Layout Randomisation~\cite{pax-aslr},
 heap cookies~\cite{heapcookies} or data execution prevention try to make this harder for
 an attacker. Attackers have developed techniques such as heap spraying~\cite{heapspray} and heap
 feng shui~\cite{fengshui} to help bypass those mitigations.

Nail parsers use two arena allocators~\cite{arena}, one for the temporary syntax tree and one for the
result. If Nail were to use the normal system allocator, the attacker could cause two packets to be
parsed at the same time (or, given a sufficiently complicated input, two structure within the same
grammar to be parsed sequentially) such that the (invalid) intermediate results from one parse,  
 If there is a buffer overflow past the
end of one input, it is much harder for the attacker craft a sequence of packets to trick the allocator into
placing a known value next to that input.
